---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Søren Orm Hansen"
date: "October XX, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Loading packages
library(pacman)

p_load(lme4, tidyverse, caret, pROC, Metrics)

# Setting theme minimal
theme_set(theme_minimal())
```


## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

```{r}
# Load data
df <- read.csv("npause_fixed.csv")

# Fixing categories
df <- df %>%
  mutate(
    Diagnosis = as.factor(Diagnosis.y),
    pause_dur = as.numeric(pause_dur),
    Participant = as.factor(Participant),
    maybe = as.numeric(Participant) # The participant column is behaving odd later - 'maybe' was a shot in the dark that hit the target ;)
  )

```


We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

```{r}
# Creating logit to probability function
logit2prob <- function(logit) {
  odds <- exp(logit)
  probs <- odds / (1 + odds)
  return(probs)
}

```

```{r}
# Creating function for folding and for creating and saving values we are interested in
folding <- function(syntax, random) {
 k = 10

# Splitting the dataframe in the two categories to ensure that we have both categories in all folds
  control <- df %>% filter(Diagnosis == "Control") %>% 
    mutate(maybe = as.factor(maybe))
  schizophrenia <- df %>% filter(Diagnosis == "Schizophrenia") %>% 
    mutate(maybe = as.factor(maybe))

# Creating folds for the two categories
  foldsCon <-
    createFolds(
      unique(control$maybe),
      k = k,
      list = TRUE,
      returnTrain = FALSE)
  foldsSchiz <-
    createFolds(
      unique(schizophrenia$maybe),
      k = k,
      list = TRUE,
      returnTrain = FALSE)

# Creating various place holders
  train_auc = rep(NA, k)
  test_auc = rep(NA, k)
  
  train_sens = rep(NA, k)
  test_sens = rep(NA, k)
  
  train_spec = rep(NA, k)
  test_spec = rep(NA, k)

  train_PPV = rep(NA, k)
  test_PPV = rep(NA, k)
  
  train_NPV = rep(NA, k)
  test_NPV = rep(NA, k)

# Looping through the different folds
  for (i in 1:k) {
    i <- as.numeric(i)
# Creating test and train subset for the i'th fold. We add 88 because that's how many unique maybes there are in the control subset 
    train <- rbind(subset(control,!(maybe %in% foldsCon[[i]])),
                   subset(schizophrenia,!(maybe %in% (foldsSchiz[[i]] + 88))))
    test <- rbind(subset(control, maybe %in% foldsCon[[i]]),
                  subset(schizophrenia, maybe %in% (foldsSchiz[[i]] + 88)))
   
# Building models on the training set. The function do either glm or glmer models
    if(random == F) {
      model <- glm(syntax, data = train, family = "binomial")
    } else {
      model <- glmer(syntax, data = train, family = "binomial")
    }

# Creating predictions
    test <- test %>%
      mutate(prediction = logit2prob(predict(model, newdata = test, allow.new.levels = TRUE)),
             pred = as.factor(ifelse(prediction > .5, "Schizophrenia", "Control")))
    train <- train %>%
      mutate(prediction = logit2prob(predict(model)),
             pred = as.factor(ifelse(prediction > .5, "Schizophrenia", "Control")))

# Checking the quality of our predictions - if there is only one category, it will mess stuf up later
    qualiCheck <- ifelse(nlevels(test$pred) == 1, 0, 1)

# ROC for train
    train_rocCurve = roc(response = train$Diagnosis,
                         predictor = train$prediction)

# Using the quality check to avoid errors in ROC etc. for test
    if(qualiCheck == 1){
      test_rocCurve = roc(response = test$Diagnosis, predictor = test$prediction)
      test_auc[i] = as.numeric(pROC::auc(test_rocCurve))
      test_sens[i]= sensitivity(
        data = test$pred,
        reference = test$Diagnosis,
        positive = "Schizophrenia")
      test_spec[i] = specificity(
        data = test$pred,
        reference = test$Diagnosis,
        negative = "Control")
      test_PPV[i] = posPredValue(
        data = test$pred,
        reference = test$Diagnosis,
       positive = "Schizophrenia")
      test_NPV[i] = negPredValue(
      data = test$pred,
      reference = test$Diagnosis,
      negative = "Control")
      
    } else {
      test_rocCurve = 0
      test_auc[i] = 0
      test_sens[i]  = 0
      test_spec[i] = 0
      test_PPV[i] = 0
      test_NPV[i] = 0
    }

# AUC for training set
    train_auc[i] = as.numeric(pROC::auc(train_rocCurve))
    
# Sensitivity for test and train using quality check to avoid errors
    train_sens[i] = sensitivity(
      data = train$pred,
      reference = train$Diagnosis,
      positive = "Schizophrenia")
    
# Specificity for test and train using quality check to avoid errors
    train_spec[i] = specificity(
      data = train$pred,
      reference = train$Diagnosis,
      negative = "Control")

# PPV for test and train using quality check to avoid errors
    train_PPV[i] = posPredValue(
      data = train$pred,
      reference = train$Diagnosis,
      positive = "Schizophrenia")
  
# NPV for test and train using quality check to avoid errors
    train_NPV[i] = negPredValue(
      data = train$pred,
      reference = train$Diagnosis,
      negative = "Control")
    
  }
  
# Finding the mean for the different values
  train_auc <- mean(train_auc)
  test_auc <- mean(test_auc)
  train_sens <- mean(train_sens)
  train_sens <- mean(train_sens)
  test_sens <- mean(test_sens)
  train_spec <- mean(train_spec)
  test_spec <- mean(test_spec)

# Returning the values
  return(
    tibble(
      train_auc,
      test_auc,
      train_sens,
      test_sens,
      train_spec,
      test_spec
    )
  )
}
  
```

```{r}
syntax1 <- 'Diagnosis ~ IQR_scaled'
folding(syntax1, random = F)

syntax2 <- 'Diagnosis ~ IQR_scaled + (1|Participant)'
folding(syntax2, random = T)
```



```{r}
# Speech rate model (glm)
m_sr <-
  glm(Diagnosis ~ SpeechRate_scaled,
      train,
      family = "binomial")

# Creating predictions on training data
train <- train %>%
  mutate(pred = logit2prob(predict(m_sr)),
         pred_class = as.factor(ifelse(pred > .5, "Schizophrenia", "Control")))

# Creating confusion matrix on training data
confusionMatrix (
  data = train$pred_class,
  reference = train$Diagnosis,
  positive = "Schizophrenia"
)

# Creating predictions on test data
test <- test %>%
  mutate(pred = logit2prob(predict(m_sr, newdata = test)),
         pred_class = as.factor(ifelse(pred > .5, "Schizophrenia", "Control")))

# Creating confusion matrix on test data
confusionMatrix (
  data = test$pred_class,
  reference = test$Diagnosis,
  positive = "Schizophrenia"
)

# Plotting predictions
test %>% ggplot(aes(x = pause_dur, y = pred_class, color = Diagnosis)) +
  geom_jitter()

posPredValue(
      data = train$pred_class,
      reference = train$Diagnosis,
      positive = "Schizophrenia"
    )

```

```{r}
# Pause duration model
pausedur <-
  glm(Diagnosis ~ pause_dur,
        family = "binomial",
        data = df)

summary(pausedur)

# Creating predictions on training data
df1 <- df %>% 
  mutate(pred = logit2prob(predict(pausedur)),
         pred_class = as.factor(ifelse(pred > .5, "Schizophrenia", "Control"))
         )

# Creating confusion matrix on training data
confusionMatrix (data = df1$pred_class,
                 reference = df1$Diagnosis,
                 positive = "Schizophrenia")

# Plotting predictions
df1 %>% ggplot(aes(x = pause_dur, y = pred_class, color = Diagnosis.y)) + 
  geom_jitter()
```


```{r}
# ROC curve
rocCurve <- roc(response = train$Diagnosis, predictor = train$pred)
auc(rocCurve)
ci (rocCurve)
plot(rocCurve, legacy.axes = TRUE) 

```

```{r}
# IQR model
IQRscale <-
  glmer(Diagnosis ~ IQR_scaled + (1 | Participant),
        family = "binomial",
        control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE),
        data = train)

summary(pausedur)

# Creating predictions on training data
train <- train %>% 
  mutate(pred = logit2prob(predict(IQRscale)),
         pred_class = as.factor(ifelse(pred > .5, "Schizophrenia", "Control"))
         )

# Creating confusion matrix on training data
confusionMatrix (data = train$pred_class,
                 reference = train$Diagnosis,
                 positive = "Schizophrenia")

# Creating predictions on test data
test <- test %>% 
  mutate(pred = logit2prob(predict(IQRscale, newdata = test, allow.new.levels = TRUE)),
         pred_class = as.factor(ifelse(pred > .5, "Schizophrenia", "Control"))
         )

# Creating confusion matrix on test data
confusionMatrix (data = test$pred_class,
                 reference = test$Diagnosis,
                 positive = "Schizophrenia")

# Plotting predictions
train %>% ggplot(aes(x = IQR_scaled, y = pred, color = Diagnosis.y)) + 
  geom_point()
```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidy-models set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)

```{r}
```
